Necessity is the mother of genetic invention

### 2017-11-02 ms.tex
migration model + MWF comments

% Title page
\title{Necessity is the mother of invention}

@@

% Abstract
\begin{abstract}
Generation of variation is considered detrimental in well-adapted
populations evolving under constant selection.
Therefore, the rate at which variation is generated through processes such as mutation,
migration, and innovation is reduced by natural selection.
However, departures from this \emph{reduction principle} have been demonstrated.
Here we analyze a general model of evolution under constant selection where the rate at which variation is generated can depend on the individual fitness.
We find a general and simple result:
if individuals with below-average fitness increase the rate at which variation is generated,
then the population mean fitness increases.
This principle applies to phenomena such as stress-induced mutagenesis,
condition-dependent dispersal and outcrossing,
and even to the famous proverb: \emph{"Necessity is the mother of invention"}.
\end{abstract}

@@

According to the \emph{reduction principle}, in populations that evolve near an equilibrium between natural selection and a process that generates variation (i.e. mutation, migration, or recombination), selection favors modifiers that decrease the rate of variation generation.

@@

The assumptions and conditions for the validity of the \emph{unified reduction principle} are~\citep{Altenberg2017}:
(i) effectively infinite population size, (ii) constant-viability selection, (iii) a population at an equilibrium, and (iv) \emph{linear variation} -- the equal scaling of transition probabilities by the modifier.
A departure from the latter assumption occurs if two variation producing processes interact~\citep{Feldman1980,Altenberg2012}.
Departures from the \emph{reduction principle} have also been demonstrated when conditions (i)-(iii) are not met, see for example~\citet{Holsinger1986} and references therein.

Another departure from the \emph{linear variation} assumption of the \emph{reduction principle} for mutation rates involves a mechanism commonly called \emph{stress-induced mutagenesis}, in which the mutation rate increases in individuals with poor fitness.
\citet{Ram2012} demonstrated that modifiers that increase the mutation rate of individuals with below-average fitness actually \textbf{increase} the population mean fitness, rather than decrease it.
Their analysis included an infinite population size and constant-viability selection, such that the only departure from the \emph{reduction principle} assumptions is due to the differential effect the modifier has on different genotypes.
\citet{Ram2012} stated that their result represents a departure from the \emph{reduction principle}, but did not explain this departure.
Their analysis was specific to a model that classified individuals by the number of mutant genes in their genotype, similar to~\citet{Kimura1966} and~\citet{Haigh1978}.

Here, we present a general evolutionary model where the rate at which variation is generated depend on the type of individual, and specifically on the its fitness.
Our result shows that low fitness individuals should produce more variation then high fitness individuals.

@@

\subsection*{General model.}
Consider a large haploid population with an arbitrary set of types $A_1, A_2, ...$.
The change in the frequencies of these types is described by the transformation $f \to f'$: 
\begin{equation}\label{eq:model}
\bar{\omega} f' = (I-C+MC)D f,
\end{equation}
where $f_k$ is the frequency of type $A_k$, such that $f$ is a stochastic vector ($\sum_k{f_k} = 1$);
$C_k$ is the probability that an individual of type $A_k$ will transition to another type ($C$ is a diagonal non-negative matrix with diagonal elements $C_k$);
$M_{i,j}$ is the transition probability from type $A_j$ to type $A_i$ given that an individual of type $A_j$ is transitioning ($M$ is a stochastic matrix: $\sum_j {M_{i,j}} = 1$ for any column \emph{j}); 
$D_k$ is the relative fitness of type $A_k$ ($D$ is a diagonal matrix with diagonal elements $0 < D_k \le 1$; fitness values are relative to the fitness of the fittest type); 
$I$ is the identity matrix;
and $\bar{\omega}$ is a normalizing factor such that $\sum_k{f'_k}=1$, which turns out to be the population mean fitness (see~\nameref{sec:AppA}).

The types $A_k$ can be encoded by a single or multiple genetic loci or non-genetic traits.
Importantly, type transmission is vertical and uni-parental (the type is transmitted from a single parent to the offspring), independent of the frequencies of the other types, and is fully described by the matrix $M$. Specifically, transmission infidelity causes type transitions.
Therefore, this model precludes processes such as recombination, social learning, sexual outcrossing, and any horizontal or oblique transmission, as these processes are frequency-dependent~\cite[pg.~54]{Cavalli-Sforza1981}.

Transition between types is determined by a combination of two effects:
(i) the probability to transition \emph{out} of type $A_k$ is determined by $C_k$;
(ii) given a transition out of type $A_i$, the distribution of the destination types $A_j$ is given by $M_{j,i}$ (note the index order).
We focus on a modifier that has no direct effect on fitness and whose sole function is to determine the rate of transition out of the different types ($C_k$).
Importantly, this modifier can determine different rates for different types, therefore $C$ is a diagonal matrix rather than a scalar. The case of a scalar in covered by the \emph{reduction principle}, see~\citet{Altenberg2017}.

@@

\subsection*{Mutation modification model 1.}
Here we consider a large population of haploids and a trait determined by a major locus with $n$ possible alleles $A_1, A_2, ..., A_n$.
Linked to to this locus is a modifier locus that determines the mutation rates $\mu_k$ of individuals with allele $A_k$.
Specifically, with probability $1-mu_k$, the allele $A_k$ does not mutate, and with probability $\frac{\mu_k}{n-1}$, the allele $A_k$ mutates to $A_j$ for any $j \ne k$.
This is an extension of a model presented by~\citet{Altenberg2017}, allowing for the mutation rate of an individual $\mu_k$ to depend on the identity of the allele $A_k$.

Let the frequency of $A_k$ in the present generation be $f_k$ such that $f_k \ge 0$, $\sum_k{f_k}=1$, $\vec{f}=(f_1, \ldots, f_n)\tr$.
Then after selection and mutation $\vec{f'}$ in the next generation is given by
\begin{equation}
f_k' = (1 - \mu_k) \frac{w_k}{\bar{w}} f_k + \frac{1}{n-1} \sum_{k \ne j}{\mu_j \frac{w_j}{\bar{w}} x_j},
\label{eq:mutation_model_1}
\end{equation}
for all $k=1,2,\ldots,n$, where $\bar{w} = \sum_{k=1}^{n}{w_k f_k}$ is the population mean fitness.

@@

Here, we consider a large population of haploids in which individuals with genotype $A_k$ are characterized by the number of deleterious or mutant alleles in their genotype, $k$.
A modifier locus then determines the probability that a mutation occurs in individuals with $k$ mutant alleles, $m_k$.
In this case, if we assume a low mutation rate $m_k$ such that two mutations are unlikely to occur in the same individual in one generation.
When a mutation occurs it is \emph{deleterious} with probability $\delta$, generating a new mutant allele and moving the individual from $A_k$ to $A_{k+1}$, or it is beneficial with probability $\beta$, moving the individual from $A_k$ to $A_{k-1}$.
Note that such beneficial mutations can be either compensatory or back-mutations, and that mutations are neutral with probability $1-\delta-\beta$.
This model has been presented and analyzed by~\citet{Ram2012}.

Let the frequency of $A_k$ in the present generation be $f_k$ such that $f_k \ge 0$, $\sum_k{f_k}=1$, $\vec{f}=(f_1, \ldots, f_n)\tr$.
Then after selection and mutation $\vec{f'}$ in the next generation is given by
\begin{equation}
f_0' = (1 - \delta m_0) \frac{w_0}{\bar{w}} f_0 + \beta m_{1} \frac{w_{1}}{\bar{w}} f_{1},
\end{equation}
and
\begin{multline}
f_k' = (1 - m_k(\beta+\delta)) \frac{w_k}{\bar{w}} f_k + \\
\delta m_{k-1} \frac{w_{k-1}}{\bar{w}} f_{k-1} + \beta m_{k+1} \frac{w_{k+1}}{\bar{w}} f_{k+1},
\label{eq:mutation_model_2}
\end{multline}
for all $k=1,2,\ldots$, where $w_k$ is the fitness of an individual with $k$ mutant alleles and $\bar{w} = \sum_{k \ge 0}{w_k f_k}$ is the population mean fitness.

@@

\begin{align}
C_k = m_k, \quad
D_k = w_k, \quad
M_{i,j} = \begin{cases}
\beta, & i = j-1 \\
1-\beta-\delta, & i = j \\
\delta, & i = j+1 \\
1-\delta, & i=j=0 \\
0, & \text{otherwise}
\end{cases}
\end{align}
so that Eq.~\ref{eq:mutation_model_2} is equivalent to Eq.~\ref{eq:model}.

@@

@@

% Mean fitness principle
The equilibrium of the frequency vector $f$ in Eq.~\ref{eq:model} is given by the equilibrium frequency vector $v$ and the stable population mean fitness
$\bar{\omega}^*$, namely
\begin{equation}\label{eq:model_equilibrium}
\bar{\omega}^* v = (I-C+MC)D v
\end{equation}

@@

\begin{mfp}
Let $\bar{\omega}^*$ be the leading eigenvalue of $(I-C+MC)D$, $u$ and $v$ be the corresponding left and right eigenvectors, and $[M]_k$ the $k$-th column of $M$. Then
\begin{equation}\label{eq:theorem}
\frac{\partial \bar{\omega}^*}{\partial C_k} = 
\frac{u_k v_k}{C_k} (\bar{\omega}^* - D_k),
\end{equation}
or in simpler terms,
\begin{equation}
sign\frac{\partial \bar{\omega}^*}{\partial C_k} = 
sign(\bar{\omega}^* - D_k), 
\end{equation}
and therefore increased transition out of type $k$ will increase the stable population mean fitness if and only if the fitness of type $k$ is below the stable population mean fitness.
\end{mfp}

@@

The corresponding equation to Eq.~\ref{eq:model_equilibrium} for the left
eigenvector $u$ is 
\begin{equation}
u\tr \bar{\omega}^* = u\tr (I - C + MC) D,
\end{equation}
which gives us a relation between $\bar{\omega}^*$ and the $k$
element of $u$:
\begin{equation}
u_k \bar{\omega}^* = (1-C_k) D_k u_k + C_k D_k u\tr [M]_k.
\end{equation}
Multiplying both sides by $v_k$ (which is positive if $D_k>0$) and rearranging, we get 
\begin{equation}
\frac{u_k v_k}{C_k} (\bar{\omega}^* - D_k) = v_k D_k(u\tr [M]_k - u_k).
\end{equation}
Plugging this in Eq.~\ref{eq:theorem_intermediate} yields:
\begin{equation}
\frac{\partial \bar{\omega}^*}{\partial C_k} = 
\frac{u_k v_k}{C_k} (\bar{\omega}^* - D_k),
\end{equation}
which completes the proof.
\end{proof}

Note that if the modifier has the same effect on all types (i.e.,
$C_k = \alpha$ for all $k$) then Eq.~\ref{eq:model_equilibrium}
becomes $\bar{\omega}^* v = ((1-\alpha)I + \alpha M)Dv$,
and the \emph{unified reduction principle} is in effect~\cite[eqs. 65, 72]{Altenberg2017}.
Therefore, we can expect modifiers that reduce $\alpha$ to be favored by natural selection.
Indeed, setting $C = \alpha I$ in Eq.~\ref{eq:model_equilibrium} and proceeding as in the proof, we get
$$
\frac{\partial \bar{\omega}^*}{\partial \alpha} = 
\bar{\omega}^* - 1 \le 0,
$$
because $\bar{\omega}^* = \sum_i{v_i D_i}$ (see~\nameref{sec:AppA}) and $D_i \le 1$ for all $i$.

>>>>

% Title page
\title{Necessity is the mother of genetic invention}

@@

% Abstract
\begin{abstract}
Generation of variation is considered detrimental in well-adapted populations evolving under constant selection.
Therefore, in a constant environment, genetic modifiers that reduce the rate at which variation is generated by processes such as mutation and migration, succeed.
However, departures from this \emph{reduction principle} have been demonstrated.
Here we analyze a general model of evolution under constant selection where the rate at which variation is generated can depend on the fitness of the individual.
We find a general and simple result:
if individuals with below-average fitness increase the rate at which variation is generated, then the population mean fitness increases.
This principle applies to phenomena such as stress-induced mutagenesis, condition-dependent dispersal and outcrossing,
and even to the famous proverb: \emph{"Necessity is the mother of genetic invention"}.
\end{abstract}

@@

According to the \emph{reduction principle}, in populations that evolve near an equilibrium between natural selection and a process that generates variation (i.e. mutation, migration, or recombination), selection favors neutral modifiers that decrease the rate at which variation is generated.

@@

The assumptions and conditions for the validity of the \emph{unified reduction principle} by neutral genetic modifiers are~\citep{Altenberg2017}:
(i) effectively infinite population size, (ii) constant-viability selection, (iii) a population at an equilibrium, and (iv) \emph{linear variation} -- the equal scaling of transition probabilities by the modifier.
A departure from the latter assumption occurs if two variation producing processes interact~\citep{Feldman1980,Altenberg2012}.
Departures from the \emph{reduction principle} have also been demonstrated when conditions (i)-(iii) are not met, see for example~\citet{Holsinger1986} and references therein.

Another departure from the \emph{linear variation} assumption of the \emph{reduction principle} for mutation rates involves a mechanism commonly called \emph{stress-induced mutagenesis}, in which the mutation rate increases in individuals with poor fitness.
\citet{Ram2012} demonstrated that modifiers that increase the mutation rate of individuals with below-average fitness actually \textbf{increase} the population mean fitness, rather than decrease it.
Their analysis assumed an infinite population size and constant-viability selection, such that the only departure from the \emph{reduction principle} assumptions is due to the differential effect the modifier has on different genotypes.
\citet{Ram2012} stated that their result represents a departure from the \emph{reduction principle}, but did not explain this departure.
Their analysis was specific to a model that classified individuals by the number of mutant genes in their genotype, similar to those of~\citet{Kimura1966} and~\citet{Haigh1978}.

Here, we present a general evolutionary model where the rate at which variation is generated depends on the type of individual, and specifically on the its fitness.
Our result shows that low fitness individuals should produce more variation than high fitness individuals.

@@

\subsection*{General model.}
Consider a large haploid population with an arbitrary (possibly infinite) set of types $A_1, A_2, \ldots$.
The change in the frequencies of these types is described by the transformation $f \to f'$: 
\begin{equation}\label{eq:model}
\bar{\omega} f' = (I-C+MC)D f,
\end{equation}
where $f_k$ is the frequency of type $A_k$ ($f=(f_1, f_2, \ldots)$ is a stochastic vector: $\sum_k{f_k} = 1$);
$C_k$ is the probability that an individual of type $A_k$ will transition to another type ($C$ is a diagonal non-negative matrix with diagonal elements $C_k$);
$M_{i,j}$ is the transition probability from type $A_j$ to type $A_i$ given that an individual of type $A_j$ is transitioning ($M$ is a column-stochastic matrix: $\sum_j {M_{i,j}} = 1$ for any column \emph{j}); 
$D_k$ is the relative fitness of type $A_k$ ($D$ is a diagonal matrix with diagonal elements $D_k$, where we can assume $0 < D_k \le 1$, i.e., fitness values are relative to the fitness of the fittest type); 
$I$ is the identity matrix;
and $\bar{\omega}$ is a normalizing factor such that $\sum_k{f'_k}=1$ and is equal to the population mean fitness (see~\nameref{sec:AppA}).

The types $A_k$ can represent a single or multiple genetic loci or non-genetic traits.
Importantly, type transmission is vertical and uni-parental (the type is transmitted from a single parent to the offspring), independent of the frequencies of the other types, and is fully described by the matrix $M$. Specifically, transmission infidelity causes type transitions.
Therefore, this model precludes processes such as recombination, social learning, sexual outcrossing, and any horizontal or oblique transmission, as these processes are frequency-dependent~\cite[pg.~54]{Cavalli-Sforza1981}.

Transition between types is determined by a combination of two effects:
(i) the probability to transition \emph{out} of type $A_k$ is determined by $C_k$;
(ii) given a transition out of type $A_i$, the distribution of the destination types $A_j$ is given by $M_{j,i}$ (note the index order).
We focus on a modifier that has no direct effect on fitness and whose sole function is to determine the rate of transition out of the different types ($C_k$).
Importantly, this modifier can determine different rates for different types; that is, $C_j \ne C_i$ for some $i,j$. The case $C_j = C_j$ for all $i,j$ is covered by the \emph{reduction principle}, see~\citet{Altenberg2017}.

@@

\subsection*{Mutation modification model 1.}

Here we consider a large population of haploids and a trait determined by a major locus with $n$ possible alleles $A_1, A_2, ..., A_n$.
Linked to to this locus is a modifier locus that determines the mutation rates $\mu_k$ of individuals with allele $A_k$.
Specifically, with probability $1-\mu_k$, the allele $A_k$ does not mutate, and with probability $\frac{\mu_k}{n-1}$, the allele $A_k$ mutates to $A_j$ for any $j \ne k$.
This is an extension of a model studied by~\citet{Altenberg2017}, that allows for the mutation rate of an individual, $\mu_k$, to depend on the identity of the allele $A_k$.

Let the frequency of $A_k$ in the present generation be $f_k$ with $f_k \ge 0$ and $\sum_k{f_k}=1$.
Then after selection and mutation, $f'_k$ in the next generation is given by
\begin{equation}
f_k' = (1 - \mu_k) \frac{w_k}{\bar{w}} f_k + \frac{1}{n-1} \sum_{k \ne j}{\mu_j \frac{w_j}{\bar{w}} x_j},
\label{eq:mutation_model_1}
\end{equation}
for all $k=1,2,\ldots,n$, where $\bar{w} = \sum_{k=1}^{n}{w_k f_k}$ is the population mean fitness.

@@

Here, we consider a large population of haploids in which individuals with genotype $A_k$ are characterized by the number $k$ of deleterious or mutant alleles in their genotype.
A modifier locus then determines the probability $m_k$ that a mutation occurs in individuals with $k$ mutant alleles.
In this case, we assume a low mutation rate $m_k$ such that two mutations are unlikely to occur in the same individual in one generation.
When a mutation occurs it is \emph{deleterious} with probability $\delta$, generating a new mutant allele and moving the individual from $A_k$ to $A_{k+1}$, or it is beneficial with probability $\beta$, moving the individual from $A_k$ to $A_{k-1}$.
Note that such beneficial mutations can be either compensatory or back-mutations, and that mutations are neutral with probability $1-\delta-\beta$.
This model has been analyzed by~\citet{Ram2012}.

Let the frequency of $A_k$ in the present generation be $f_k$ such that $f_k \ge 0$ and $\sum_k{f_k}=1$.
Then after selection and mutation $\vec{f'}$ in the next generation is given by
\begin{equation}
f_0' = (1 - \delta m_0) \frac{w_0}{\bar{w}} f_0 + \beta m_{1} \frac{w_{1}}{\bar{w}} f_{1},
\end{equation}
and
\begin{multline}
f_k' = (1 - m_k(\delta+\beta)) \frac{w_k}{\bar{w}} f_k + \\
\delta m_{k-1} \frac{w_{k-1}}{\bar{w}} f_{k-1} + \beta m_{k+1} \frac{w_{k+1}}{\bar{w}} f_{k+1},
\label{eq:mutation_model_2}
\end{multline}
for all $k=1,2,\ldots$, where $w_k$ is the fitness of an individual with $k$ mutant alleles and $\bar{w} = \sum_{k \ge 0}{w_k f_k}$ is the population mean fitness.

@@

\begin{equation}
C_k = m_k, \quad
D_k = w_k,
%M_{i,j} = \begin{cases}
%\beta, & i = j-1 \\
%1-\beta-\delta, & i = j \\
%\delta, & i = j+1 \\
%1-\delta, & i=j=0 \\
%0, & \text{otherwise}
%\end{cases}
\end{equation}
and
\begin{equation}
M_{i,j} = \begin{bmatrix}
1-\delta & \beta & 0 &  \cdots \\
\delta & 1-\delta-\beta & \beta &  \ddots \\
0 & \delta & 1-\delta-\beta & \ddots \\
\vdots & \ddots & \ddots & \ddots 
\end{bmatrix}
\end{equation}
so that Eq.~\ref{eq:mutation_model_2} is equivalent to Eq.~\ref{eq:model}.

@@

Here we assume a large population of haploids that occupies two demes and a modifier locus that determine the migration rates between the demes, $m_k$. 
Let the fitness of genotypes $A_k$ be $w_k$ in deme 1 and $u_k$ in deme 2 for $k=1,2,\ldots,n$.
Let the frequencies of $A_k$ in deme 1 be $x_k$ and in deme 2 $y_k$ for $k=1,2,\ldots,n$ with $x_k \ge 0$, $y_k \ge 0$, $\sum_{k=1}^{n}{x_k} = 1$. and $\sum_{k=1}^{n}{y_k} = 1$.

Then following selection and migration, after one generation the new frequencies $x_k'$ and $y_k'$ are given by
\begin{align}
x_k' = (1-m_k) \frac{w_k}{\bar{w}} x_k + m_k \frac{u_k}{\bar{u}} y_k \\
y_k' = m_k \frac{w_k}{\bar{w}} x_k + (1 - m_k) \frac{u_k}{\bar{u}} y_k
\end{align}
for $k=1,2,\ldots,n$, where $\bar{w} = \sum_{k \ge 0}{w_k x_k}$ and $\bar{u} = \sum_{k \ge 0}{u_k y_k}$ is the population mean fitness in deme 1 and 2, respectively.
This is an extension of a model presented by~\citet{Altenberg2017}, allowing for the migration rate of an individual, $m_k$, to depend on the identity of the allele $A_k$.

To rephrase this model in terms of the general model, we define
\begin{equation}
\begin{aligned}
f = (x_1, \ldots, x_n, y_1, \ldots, y_n)\tr, \\
C = diag[m_1, \ldots, m_n, m_1, \ldots, m_n], \\
D = diag[w_1, \ldots, w_n, u_1, \ldots, u_n], \\
M = \begin{bmatrix}
0 & I_n \\
I_n & 0
\end{bmatrix}
\end{aligned}
\end{equation}
where $I_n$ is the n-by-n identity matrix.

@@

% Mean fitness principle
With the equilibrium of the frequency vector $f$ in Eq.~\ref{eq:model} as $v$ and the stable population mean fitness as $\bar{\omega}^*$, then
\begin{equation}\label{eq:model_equilibrium}
\bar{\omega}^* v = (I-C+MC)D v
\end{equation}

@@

\begin{mfp}
Let $\bar{\omega}^*$ be the leading eigenvalue of $(I-C+MC)D$, $u$ and $v$ be the corresponding left and right eigenvectors, and $[M]_k$ the $k$-th column of $M$. Then
\begin{equation}\label{eq:theorem}
\frac{\partial \bar{\omega}^*}{\partial C_k} = 
\frac{u_k v_k}{C_k} (\bar{\omega}^* - D_k),
\end{equation}
or in simpler terms,
\begin{equation}
sign\frac{\partial \bar{\omega}^*}{\partial C_k} = 
sign(\bar{\omega}^* - D_k), 
\end{equation}
and therefore increased transition from type $k$ will increase the stable population mean fitness if and only if the fitness of type $k$ is below the stable population mean fitness.
\end{mfp}

@@

The corresponding equation to Eq.~\ref{eq:model_equilibrium} for the left
eigenvector $u$ is 
\begin{equation}
u\tr \bar{\omega}^* = u\tr (I - C + MC) D,
\end{equation}
which gives us a relation between $\bar{\omega}^*$ and the $k$
element of $u$:
\begin{equation}
u_k \bar{\omega}^* = (1-C_k) D_k u_k + C_k D_k u\tr [M]_k.
\end{equation}
Multiplying both sides by $v_k$ (which is positive if $D_k>0$) and rearranging, we get 
\begin{equation}
\frac{u_k v_k}{C_k} (\bar{\omega}^* - D_k) = v_k D_k(u\tr [M]_k - u_k).
\end{equation}
Plugging this in Eq.~\ref{eq:theorem_intermediate} yields:
\begin{equation}
\frac{\partial \bar{\omega}^*}{\partial C_k} = 
\frac{u_k v_k}{C_k} (\bar{\omega}^* - D_k).
\end{equation}
Finally, note that $u_k, v_k, C_k \ge 0$ and therefore
\begin{equation}
sign \frac{\partial \bar{\omega}^*}{\partial C_k} = 
sign (\bar{\omega}^* - D_k),
\end{equation}
which completes the proof.
\end{proof}

Note that if the modifier has the same effect on all types (i.e.,
$C_k = \alpha$ for all $k$) then Eq.~\ref{eq:model_equilibrium}
becomes $\bar{\omega}^* v = ((1-\alpha)I + \alpha M)Dv$,
and the \emph{unified reduction principle} is in effect~\cite[eqs. 65, 72]{Altenberg2017}.
Therefore, we can expect modifiers that reduce $\alpha$ to be favored by natural selection.
Indeed, setting $C = \alpha I$ in Eq.~\ref{eq:model_equilibrium} and proceeding as in the above proof, we get
$$
\frac{\partial \bar{\omega}^*}{\partial \alpha} = 
\bar{\omega}^* - 1 \le 0,
$$
because $\bar{\omega}^* = \sum_i{v_i D_i}$ (see~\nameref{sec:AppA}) and $D_i \le 1$ for all $i$.
